\documentclass[a4paper,12pt]{article}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage[english]{babel}	% deutsche Umgebung (neue Rechtschreibung)
\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}	
\usepackage{pdfpages}
\usepackage{times} 		% Times als Schrift
\usepackage{setspace}		% Hilfspaket fr Satzspiegelberechnung
\usepackage{remreset}		% Ermglicht die Rckname von internen Redefinitionen durch subpakete
\usepackage{xspace}		% "intelligente" Leerzeichen nach Zitaten
\usepackage{listings}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{bbm}
\usepackage{soul}
\usepackage{amssymb}
\usepackage{ulem}
\usepackage{natbib}
\usepackage{layout}		% zu Testzwecken
\usepackage{eurosym} %für €
\usepackage[noindentafter]{titlesec} %dass paragraphs auch nummeriert werdn können
\usepackage{lscape}
\usepackage[none]{hyphenat}
\newtheorem{hypothesis}{Hypothesis}
\sloppy
\usepackage{float}
\usepackage{array} % für extra zeilenhöhe
\usepackage{color}
\usepackage{multirow}
\usepackage{lscape}
\usepackage{arydshln}
\usepackage[a4paper]{geometry}
\usepackage{booktabs}

\textwidth 16cm
\topmargin -8mm
\oddsidemargin 8mm
\evensidemargin -8mm
\textheight 24cm
\topskip 0cm
\onehalfspacing
\parindent 0cm
\parskip 1ex
\footnotesep .4cm
\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}
\setlength{\parindent}{0pt}
\begin{document}
\begin{titlepage}

\title{Incentivizing Creativity: a Large-Scale Experiment with Tournaments and Gifts 
\footnote{We gratefully acknowledge financial support from the Germany's Federal Ministry of Education and Research framework program ``Economics of Science.'' Further, we thank the SEEK research program for their financial support as well as the many seminar and conference participants whose valuable contributions made this paper much better. Specifical thanks goes to Dirk Engelmann, Michael Kosfeld, Francois Laisney, Steve Levitt, John Morgan,  Nick Zubanov, and Thomas Zwick. We thank Christian Bommer, Michael D\"orsam, Radost Holler, Sascha Lehman, Johannes Moser, Vera Schmitz, Jan Schneiderwind, Mattie Thoma, Jeannine van Reeken, and Timo Vogelsang for outstanding research assistance. This paper was previously circulated as ``Creativity is different: Comparing rewards across a creative and a routine task'' and ``Rewards and Performance:
A Comparison Across a Creative and a Routine Task''.}}
%\normalfont
\author{Christiane Bradler\footnote{Volkswagen, Germany. E-mail: christianebradler@yahoo.de.},\ Susanne Neckermann\footnote{University of Chicago, ZEW, and CREMA. E-mail: sneckermann@uchicago.edu.},\ Arne Jonas Warnke\footnote{ZEW Centre for European Economic Research Mannheim. E-mail: warnke@zew.de.}}
\date{June 2017}
\maketitle
\thispagestyle{empty}

\begin{abstract}
\noindent
This paper reports the results from a large-scale laboratory experiment that compares 
 the impact of tournament incentives and wage gifts on a creative and a simple task. 
We find that tournaments substantially increase output in both tasks. By comparison,
wage gifts are significantly more effective in the simple task. Additional treatments suggest that it may be
the uncertain mapping between effort and output in the creative task that reduces 
reciprocity. 
Our findings provide a rationale for the frequent use of tournaments when seeking to motivate creative output. 


%While a large literature documents the effectiveness of explicit (tournaments) 
%and implicit (wage gifts) incentives for simple tasks, 
%we know little about how to foster creativity. 
%This paper presents a comprehensive assessment 
%of both types of rewards and their effects on both 
%a creative and a simple task in one unified set-up. 
%Tournaments effectively trigger performance in both tasks; there 
%is no evidence for crowding out. Interestingly, the gift 
%triggers reciprocity only in the simple task. 
%Additional treatments reveal that reciprocity is inhibited in the creative task
%because agents lack perfect information on how valuable their ideas are. 
%Policy implications are discussed.
\end{abstract}

JEL Classification: C91, D03, J33, M52\\
Keywords: creativity, incentives, tournament, reciprocity, experiment, crowding-out
\end{titlepage}

\section{Introduction}
\input{1_Introduction_150809.tex}

\section{Literature}
\input{1.5_Literature.tex}

\section{The Experiment}\label{chap:experiment}
\input{4_Experiment_150313.tex}

\section{Main Results}\label{chap:results}
\input{5_simple_and_main_analyses.tex}

%\section{Supplementary Investigations}\label{chap:suppl}
%\input{6_supplementary.tex}

\section{Conclusion}\label{chap:discussion}
\input{7_Discussion_Conclusion_150204.tex}
                 

\newpage
\addcontentsline{toc}{section}{Reference}
\bibliographystyle{elsart-harv}
\bibliography{Bib_Creativity}

\newpage
\titleformat{\section}
  {\normalfont\Large\bf\centering}
  {\thesection}
  {1em}
  {}
	\titleformat{\subsection}
  {\normalfont\large\bf\centering}
  {\thesubsection}
  {1em}
  {}
\section*{Figures and Tables}
\input{10b_Appendix_Figures_150921.tex}
\input{10a_Appendix_Tables_151212.tex}
\setcounter{section}{0}
\renewcommand*{\theHsection}{chX.\the\value{section}}
\section*{Appendix}
\subsection*{Translated Instructions (Original in German)} \label{instructions}
\input{Instructions_final.tex} 
%\input{14_KEKundEKE.tex}
%\input{15_Feedback.tex}
%\input{16_Breaks.tex}
\clearpage
\subsection*{Webappendix} \label{instructions}
\input{20_Webappendix.tex}
\end{document}

